---
title: "CapStoneProject"
author: "Àran Kieskamp"
date: "19 May 2020"
output: html_document
---

# Capstone project

#2 Exploratory Data Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyverse)
library(pander)
library(lubridate)
library(gridExtra)
library(BiocManager)
library(affy)
library(DESeq2)
library(pheatmap)
```


```{r Loading data}
data <- read.table("GSE132824_Abiotic_RNA-Seq.Readcount.txt", header = T)
data <- as_tibble(data)

head(data, n = 5L)
dim(data)
str(data)

#splitting data on different methods/*-

cold <- 1:15
heat <- 16:30
osmosis <- 31:45
control <- 46:63
saline <- 64:78

```

```{r}
summary(data)
```

3.4.1

when looking at the summary output of the data, it shows that all minimums and 1st quartiles are 0. and the 3rd quartiles and the max are really far apart.


to visualize this in a boxplot a log2 operation will be performed and the outliers will be removed. to be able to successfully use the log2, a pseudo count of 1 will be added
```{r}
boxplot(log2(data + 1), outline = F)
```
```{r}
plotDensity(log2(data + 0.1))
```
```{r}
barplot(colSums(data) / 1e6)
```
```{r}

(ddsMat <- DESeqDataSetFromMatrix(countData = data, colData = data.frame(samples = names(data)), design = ~1))
```
```{r}
rld.dds <- vst(ddsMat)
rld <- assay(rld.dds)
```

```{r}
dists <- dist(t(rld))
```

```{r}
distsMatrix <- as.matrix(dists)

annotation <- data.frame("name" =
                                 c(
                                         rep("cold", times = length(cold)),
                                         rep("heat", times = length(heat)),
                                         rep("osmosis", times = length(osmosis)),
                                         rep("control", times = length(control)),
                                         rep("saline", times = length(saline))
                                 )
)

# Set the rownames of the annotation dataframe to the sample names (required)
rownames(annotation) <- names(data)

pheatmap(distsMatrix,
         annotation_col = annotation,
         show_colnames = F,
         clustering_distance_rows = dists,
         clustering_distance_cols = dists
)
```

```{r}
library(PoiClaClu)
dds <- assay(ddsMat)
poisd <- PoissonDistance(t(dds))
PoisDistMatrix <- as.matrix(poisd$dd)

mdsPoisData <- data.frame(cmdscale(PoisDistMatrix))
names(mdsPoisData) <- c('x_coord', 'y_coord')

groups <- factor(c(
        rep("cold", times = length(cold)),
        rep("heat", times = length(heat)),
        rep("osmosis", times = length(osmosis)),
        rep("control", times = length(control)),
        rep("saline", times = length(saline))
))

coldata <- names(data)

ggplot(mdsPoisData, aes(x_coord, y_coord, color = groups, label = coldata)) +
        geom_text(size = 4) +
        ggtitle('Multi Dimensional Scaling') +
        labs(x = "Poisson Distance", y = "Poisson Distance") +
        theme_bw()
```

# chapter 4

```{r}
counts.fpm <- log2(fpm(ddsMat, robust = TRUE) + 1)

```
